---
title: "Predicting Asthma Rates in California"
author: "Pranjal Srivastava"
date: "5/8/2022"
output:
  pdf_document:
    latex_engine: xelatex
  highlight: pygments
  toc: yes
  toc_float: yes
---


```{r}
setwd("/Users/pranjalsrivastava/Desktop/Projects/Predicting Emergency Department Visits for Asthma")

library(tidyr)
library(magrittr)
library(ggplot2)
library(dplyr)
library(haven)
library(nortest)
library(leaps)
library(faraway)


```

##Data Cleaning
```{r}
cal_envir1 <- read.csv("Clean_Cal_Environtmentdata.csv")
#cal_envir1 <- cal_envir %>% dplyr::select(Pesticides, `Tox. Release`, `Pollution Burden Score`, PM2.5, Traffic, Ozone, Asthma) 
summary(cal_envir1)
cal_envir2 <- na.omit(cal_envir1)
summary(cal_envir2)
write.csv(cal_envir2, "Clean_Cal_Environtmentdata.csv")
```

##Univariate Analysis & Standard Deviation
```{r}
#Print observations
nrow(cal_envir2)
#Univariate
summary(cal_envir2)
#SD
sd(cal_envir2$Pesticides)
sd(cal_envir2$'Tox. Release')
sd(cal_envir2$'Pollution Burden Score')
sd(cal_envir2$PM2.5)
sd(cal_envir2$Traffic)
sd(cal_envir2$Ozone)
sd(cal_envir2$Asthma)

```

##Exploratory Plots
```{r}
#Histograms
par(mfrow = c(2,3))
hist(cal_envir2$Ozone, main = "Ozone", xlab = "Daily maximum 8-hour Ozone concentration" , col = "skyblue")
hist(cal_envir2$PM2.5, main = "PM 2.5" , xlab = "Annual mean PM 2.5 concentrations" , col = "skyblue")
hist(cal_envir2$Pesticides, main = "Pesticides" , xlab = "Total pounds per square mile" , col = "skyblue")
hist(cal_envir2$Traffic, main = "Traffic" , xlab = "Traffic density" , col = "skyblue")
hist(cal_envir2$Pollution.Burden.Score, main = "Pollution Burden Score" , xlab = "Pollution burden score" , col = "skyblue")
hist(cal_envir2$Tox..Release, main = "Toxic Release" , xlab = "Toxicity-weighted concentrations" , col = "skyblue")

#Scatter plots
y <- cal_envir2$Asthma
par(mfrow= c(2,3))
plot(cal_envir2$Ozone,y, 
     xlab = "Ozone", ylab = "Asthma", 
     main = "Ozone and Asthma", col = "blue")
abline(lm(y ~ cal_envir2$Ozone, data = cal_envir2), col = "red")

plot(cal_envir2$PM2.5,y, 
     xlab = "PM2.5", ylab = "Asthma", 
     main = "PM 2.5 and Asthma", col = "blue")
abline(lm(y ~ cal_envir2$PM2.5, data = cal_envir2), col = "red")

plot(cal_envir2$Pesticides,y, 
     xlab = "Pesticides", ylab = "Asthma", 
     main = "Pesticides and Asthma", col = "blue")
abline(lm(y ~ cal_envir2$Pesticides, data = cal_envir2), col = "red")

plot(cal_envir2$Traffic,y, 
     xlab = "Traffic", ylab = "Asthma", 
     main = "Traffic and Asthma", col = "blue")
abline(lm(y ~ cal_envir2$Traffic, data = cal_envir2), col = "red")

plot(cal_envir2$Pollution.Burden.Score, y, 
     xlab = "Pollution Burden Score", ylab = "Asthma", 
     main = "Pollution Burden Score and Asthma", col = "blue")
abline(lm(y ~ cal_envir2$Pollution.Burden.Score, data = cal_envir2), col = "red")

plot(cal_envir2$Tox..Release,y, 
     xlab = "Toxic Release", ylab = "Asthma", 
     main = "Toxic Release and Asthma", col = "blue")
abline(lm(y ~ cal_envir2$Tox..Release, data = cal_envir2), col = "red")
```

##Fitting a Model
```{r}
#fitting a model
model1 <- lm(Asthma ~ Ozone+PM2.5+Pesticides+Traffic+Pollution.Burden.Score+ Tox..Release, data = cal_envir2)
summary(model1)
```

##Diagnostics
```{r}
#Residuals vs. fitted plots
par (mfrow = c (1,1))
plot(fitted(model1), residuals(model1), main = "Residuals vs. Fitted" , xlab = "Fitted", ylab = "Residuals" , col = "darkblue")
abline (h=0, col = "red")

#Quantile-Quantile plots with standardized residuals
par (mfrow = c(1,2))
qqnorm(rstandard(model1), main = "Q-Q Plot" , col = "blue")
abline(0,1, col = "red")
hist (rstandard(model1), main = "Standardized Residuals", xlab =
"Standardized Residuals" , col = "blue")

#Normality Shapiro-Wilk Test
#shapiro.test(residuals(model1)) # data too large for Shapiro so using Anderson-Darling normality test

# Anderson-Darling normality test
ad.test(residuals(model1))

# Interpretation for AD test: reject the hypothesis of normality when the p-value is less than or equal to 0.05

# Outliers
range(rstudent(model1))

p <- 2 # one predictor + an intercept
n <- nrow (cal_envir2)
qt(1-.05 /(n*2),n-p-1)

#correlation
cal_envir_modified <- cal_envir2[,-1]
cal_envir_corrmatrix <- cor(cal_envir2,cal_envir_modified)
cal_envir_corrmatrix <- round(cal_envir_corrmatrix,2)
cal_envir_corrmatrix

#eigenvalues
cal_envir_eigen <- model.matrix(model1)[,-1]
e <- eigen(t(cal_envir_eigen) %*% cal_envir_eigen)
examen_e <- sqrt(e$val[1] / e$val)
sum(examen_e > 30)

#variance inflation factors
require(MASS)
require(faraway)
cal_envir_vif <- vif(model1)
cal_envir_vif
```

##Modified model Box-Cox Method
```{r}
#boxcox
a <- boxcox(model1, plotit = TRUE)

#finding lambda
a$x[which.max(a$y)]
```

##Making a new model (Model 2)
```{r}
#Model2 With y Transformed
model2 <- lm(Asthma^(10/99) ~ Ozone + PM2.5 + Pesticides + Traffic + Pollution.Burden.Score + Tox..Release, data = cal_envir2)
summary(model2)
```

##Model 2 Diagnostics
```{r}
#Residuals vs Fitted Plot 
plot(fitted(model2), rstandard(model2),
     xlab = "Fitted",
     ylab = "Residuals", col = "darkblue" , main = "Residuals vs. Fitted")
abline(h=0, col = "red")

# Quantile-Quantile plots with standardized residuals
par (mfrow = c (1,2)) 
qqnorm(rstandard(model2),main = "Q-Q Plot", col = "blue")
abline(0,1, col = "red")
hist(rstandard(model2), main = "Standardized Residuals", xlab = "Standardized Residuals" , col = "blue")
```

##Cooks Distance to Remove Outliers
```{r}
#Cooks distance
cal_envir2$cooked <- cooks.distance(model2)
sample_size <- nrow(cal_envir2)
cal_envir2$outlier <- ifelse((cal_envir2$cooked < 4/sample_size),"keep","delete")
cal_new <- cal_envir2[!(cal_envir2$outlier=="delete"),]

rname <- row.names(cal_envir2)


halfnorm(cal_envir2$cooked,4,labs = rname, ylab = "Cook's Distance" , main = "Cook's Distance vs Half-Normal")
```

##Making a New Model (Model 3) with Outliers Removed
```{r}
#Model 3 with new data, outliers removed
model3 <- lm(Asthma^(10/99) ~ Ozone + PM2.5 + Pesticides + Traffic + Pollution.Burden.Score + `Tox..Release`, data=cal_new)
summary(model3)
nrow(cal_new)
```

##Diagnostics for Model 3
```{r}
#Residuals vs Fitted Plot 
plot(fitted(model3), rstandard(model3),
     xlab = "Fitted",
     ylab = "Residuals", col = "darkblue" , main = "Residuals vs. Fitted")
abline(h=0, col = "red")

# Quantile-Quantile plots with standardized residuals
par (mfrow = c (1,2)) 
qqnorm(rstandard(model3),main = "Q-Q Plot: Model 3", col = "blue")
abline(0,1, col = "red")
hist(rstandard(model3), main = "Standardized Residuals ", xlab = "Standardized Residuals" , col = "blue")
```

##Comparing Model1 and Model3
```{r}
#QQplot M1 vs M3
par (mfrow = c (1,2)) 
qqnorm(rstandard(model3),main = "Q-Q Plot: Model 3", col = "blue")
abline(0,1, col = "red")
qqnorm(rstandard(model1),main = "Q-Q Plot: Model 1", col = "blue")
abline(0,1, col = "red")
```

##Model Selection Based on AIC
```{r}
#AIC

b <- regsubsets(Asthma^(10/99) ~ Ozone + PM2.5 + Pesticides + Traffic + Pollution.Burden.Score + Tox..Release, data=cal_new)
rs <- summary(b)
rs$which
rs$rss
n <- nrow(cal_new)
p <- 2:7
AIC <- n*log(rs$rss / n) + 2 * p
AIC
plot(AIC ~ I(p - 1), ylab = "AIC", xlab = "Number of Predictors" , col = "blue")

#Best model is model 4, with Ozone, Traffic, Pollution Burden Score, and Tox. Release as predictors
```

##Final Model (Model 4), removed predictors
```{r}
#Final Model
model4 <- lm(Asthma^(10/99) ~ Ozone + Traffic + Pollution.Burden.Score + Tox..Release, data=cal_new)
summary(model4)
confint(model4)

model4$coefficients^(99/10)

```

